---
title: "Microbiome Inhalation Laura"
output: html_notebook
---

#Load libraries
```{r include=FALSE}
library(ggplot2)
library(tidyr)
library(RColorBrewer)
library(expss)
library(dplyr)
library(BiocManager)
library(GenomicAlignments)
library(GenomicFeatures)
library(AnnotationDbi)
library(Rattus.norvegicus)
library(EnhancedVolcano)
library(TxDb.Rnorvegicus.UCSC.rn6.refGene)
library(topGO)
library(gridExtra)
library(clusterProfiler)
library(fgsea)
library(factoextra)
library(tidyverse)
library(pathview)
library(readxl)
library(stringr)
library(colorRamps)
```

```{r}
y<- read_xlsx("2023-01-25 PCB52 Inhalation 28 Day PiCRUST results.xlsx")
y<-as.data.frame(y)

```

#females
#**Dose 1: gene enrichment analysis (GSEA)**
```{r}
z<- y[order(y$FemaleCecum, decreasing = TRUE),] #reorder the genes based on significance

gene_list<-z$FemaleCecum

names(gene_list)<-z$`#NAME`
gene_list<-na.omit(gene_list)
gene_list <- gene_list[!duplicated(names(gene_list))]
head(gene_list)
```

#KEGG pathview
```{r}
kegg_organism = "ko"
kk2 <- gseKEGG(geneList     = gene_list,
               organism     = kegg_organism,
               minGSSize    = 3,
               maxGSSize    = 800,
               pvalueCutoff = 0.05,
               pAdjustMethod = "none",
               keyType       = "kegg")

data<-data.frame(kk2)
write.csv(data ,file="KEGG_results/Female_cecum_KEGG_enrichment.csv")
```

```{r}
tiff(file = 'enrichment_plots/Female_Cecum_enrichment.tiff', units="in", width=10, height=5, res=1000)
dot<-dotplot(kk2, showCategory=5, split=".sign")+
  scale_color_distiller(palette = "RdYlBu", trans= "reverse")+
  facet_grid(.~.sign)+
  ggtitle("Female Cecum KEGG Enrichment")+                                                           # Change font size
  theme(strip.text.x = element_text(size = 12))+
  theme(axis.text.y = element_text(size = 12))    

dot
dev.off()
dot
```

```{r}
# Produce the native KEGG plot (PNG)
map<- pathview(gene.data=gene_list, pathway.id="02010", species = kegg_organism)

```

#**Dose 2: gene enrichment analysis (GSEA)**
```{r}
z<- y[order(y$FemaleColon, decreasing = TRUE),] #reorder the genes based on significance

gene_list<-z$FemaleColon

names(gene_list)<-z$`#NAME`
gene_list<-na.omit(gene_list)
gene_list <- gene_list[!duplicated(names(gene_list))]
head(gene_list)
```

#KEGG pathview
```{r}
kegg_organism = "ko"
kk2 <- gseKEGG(geneList     = gene_list,
               organism     = kegg_organism,
               minGSSize    = 3,
               maxGSSize    = 800,
               pvalueCutoff = 0.05,
               pAdjustMethod = "none",
               keyType       = "kegg")

data<-data.frame(kk2)
write.csv(data ,file="KEGG_results/Female_colon_KEGG_enrichment.csv")
```

```{r}
tiff(file = 'enrichment_plots/Female_Colon_enrichment.tiff', units="in", width=10, height=5, res=1000)
dot<-dotplot(kk2, showCategory=5, split=".sign")+
  scale_color_distiller(palette = "RdYlBu", trans= "reverse")+
  facet_grid(.~.sign)+
  ggtitle("Female Colon KEGG Enrichment")+                                                           # Change font size
  theme(strip.text.x = element_text(size = 12))+
  theme(axis.text.y = element_text(size = 12))    

dot
dev.off()
dot
```

#**Dose 3: gene enrichment analysis (GSEA)**
```{r}
z<- y[order(y$FemaleIleum, decreasing = TRUE),] #reorder the genes based on significance

gene_list<-z$FemaleIleum

names(gene_list)<-z$`#NAME`
gene_list<-na.omit(gene_list)
gene_list <- gene_list[!duplicated(names(gene_list))]
head(gene_list)
```

#KEGG pathview
```{r}
kegg_organism = "ko"
kk2 <- gseKEGG(geneList     = gene_list,
               organism     = kegg_organism,
               minGSSize    = 3,
               maxGSSize    = 800,
               pvalueCutoff = 0.05,
               pAdjustMethod = "none",
               keyType       = "kegg")

data<-data.frame(kk2)
write.csv(data ,file="KEGG_results/Female_ileum_KEGG_enrichment.csv")
```

```{r}
tiff(file = 'enrichment_plots/Female_ileum_enrichment.tiff', units="in", width=10, height=5, res=1000)
dot<-dotplot(kk2, showCategory=5, split=".sign")+
  scale_color_distiller(palette = "RdYlBu", trans= "reverse")+
  facet_grid(.~.sign)+
  ggtitle("Female Ileum KEGG Enrichment")+                                                           # Change font size
  theme(strip.text.x = element_text(size = 12))+
  theme(axis.text.y = element_text(size = 12))    

dot
dev.off()
dot
```

#**Dose 4: gene enrichment analysis (GSEA)**
```{r}
z<- y[order(y$FemaleJejunum, decreasing = TRUE),] #reorder the genes based on significance

gene_list<-z$FemaleJejunum

names(gene_list)<-z$`#NAME`
gene_list<-na.omit(gene_list)
gene_list <- gene_list[!duplicated(names(gene_list))]
head(gene_list)
```

#KEGG pathview
```{r}
kegg_organism = "ko"
kk2 <- gseKEGG(geneList     = gene_list,
               organism     = kegg_organism,
               minGSSize    = 3,
               maxGSSize    = 800,
               pvalueCutoff = 0.05,
               pAdjustMethod = "none",
               keyType       = "kegg")

data<-data.frame(kk2)
write.csv(data ,file="KEGG_results/Female_Jejunum_KEGG_enrichment.csv")
```

```{r}
tiff(file = 'enrichment_plots/Female_Jejunum_enrichment.tiff', units="in", width=10, height=5, res=1000)
dot<-dotplot(kk2, showCategory=5, split=".sign")+
  scale_color_distiller(palette = "RdYlBu", trans= "reverse")+
  facet_grid(.~.sign)+
  ggtitle("Female Jejunum KEGG Enrichment")+                                                           # Change font size
  theme(strip.text.x = element_text(size = 12))+
  theme(axis.text.y = element_text(size = 12))    

dot
dev.off()
dot
```


#males

#**Dose 1: gene enrichment analysis (GSEA)**
```{r}
z<- y[order(y$MaleCecum, decreasing = TRUE),] #reorder the genes based on significance

gene_list<-z$MaleCecum

names(gene_list)<-z$`#NAME`
gene_list[gene_list==0] <- NA
gene_list<-na.omit(gene_list)
gene_list <- gene_list[!duplicated(names(gene_list))]
head(gene_list)
```

#KEGG pathview
```{r}
kegg_organism = "ko"

kk2 <- gseKEGG(geneList     = gene_list,
               organism     = kegg_organism,
               minGSSize    = 1,
               maxGSSize    = 800000,
               pvalueCutoff = 0.05,
               pAdjustMethod = "none",
               keyType       = "kegg")

data<-data.frame(kk2)
write.csv(data ,file="KEGG_results/Male_cecum_KEGG_enrichment.csv")
```

```{r}
tiff(file = 'enrichment_plots/Male_Cecum_enrichment.tiff', units="in", width=10, height=5, res=1000)
dot<-dotplot(kk2, showCategory=5, split=".sign")+
  scale_color_distiller(palette = "RdYlBu", trans= "reverse")+
  facet_grid(.~.sign)+
  ggtitle("Male Cecum KEGG Enrichment")+                                                           # Change font size
  theme(strip.text.x = element_text(size = 12))+
  theme(axis.text.y = element_text(size = 12))    

dot
dev.off()
dot
```

#**Dose 2: gene enrichment analysis (GSEA)**
```{r}
z<- y[order(y$MaleColon, decreasing = TRUE),] #reorder the genes based on significance

gene_list<-z$MaleColon

names(gene_list)<-z$`#NAME`
gene_list<-na.omit(gene_list)
gene_list <- gene_list[!duplicated(names(gene_list))]
head(gene_list)
```

#KEGG pathview
```{r}
kegg_organism = "ko"
kk2 <- gseKEGG(geneList     = gene_list,
               organism     = kegg_organism,
               minGSSize    = 3,
               maxGSSize    = 800,
               pvalueCutoff = 0.05,
               pAdjustMethod = "none",
               keyType       = "kegg")

data<-data.frame(kk2)
write.csv(data ,file="KEGG_results/Male_colon_KEGG_enrichment.csv")
```

```{r}
tiff(file = 'enrichment_plots/Male_Colon_enrichment.tiff', units="in", width=10, height=5, res=1000)
dot<-dotplot(kk2, showCategory=5, split=".sign")+
  scale_color_distiller(palette = "RdYlBu", trans= "reverse")+
  facet_grid(.~.sign)+
  ggtitle("Male Colon KEGG Enrichment")+                                                           # Change font size
  theme(strip.text.x = element_text(size = 12))+
  theme(axis.text.y = element_text(size = 12))   
dot
dev.off()
dot
```

#**Dose 3: gene enrichment analysis (GSEA)**
```{r}
z<- y[order(y$MaleIleum, decreasing = TRUE),] #reorder the genes based on significance

gene_list<-z$MaleIleum

names(gene_list)<-z$`#NAME`
gene_list<-na.omit(gene_list)
gene_list <- gene_list[!duplicated(names(gene_list))]
head(gene_list)
```

#KEGG pathview
```{r}
kegg_organism = "ko"
kk2 <- gseKEGG(geneList     = gene_list,
               organism     = kegg_organism,
               minGSSize    = 3,
               maxGSSize    = 800,
               pvalueCutoff = 0.05,
               pAdjustMethod = "none",
               keyType       = "kegg")

data<-data.frame(kk2)
write.csv(data ,file="KEGG_results/Male_ileum_KEGG_enrichment.csv")
```

```{r}
tiff(file = 'enrichment_plots/Male_ileum_enrichment.tiff', units="in", width=10, height=5, res=1000)
dot<-dotplot(kk2, showCategory=5, split=".sign")+
  scale_color_distiller(palette = "RdYlBu", trans= "reverse")+
  facet_grid(.~.sign)+
  ggtitle("Male Ileum KEGG Enrichment")+                                                           # Change font size
  theme(strip.text.x = element_text(size = 12))+
  theme(axis.text.y = element_text(size = 12))   
dot
dev.off()
dot
```

#**Dose 4: gene enrichment analysis (GSEA)**
```{r}
z<- y[order(y$MaleJejunum, decreasing = TRUE),] #reorder the genes based on significance

gene_list<-z$MaleJejunum

names(gene_list)<-z$`#NAME`
gene_list<-na.omit(gene_list)
gene_list <- gene_list[!duplicated(names(gene_list))]
head(gene_list)
```

#KEGG pathview
```{r}
kegg_organism = "ko"
kk2 <- gseKEGG(geneList     = gene_list,
               organism     = kegg_organism,
               minGSSize    = 3,
               maxGSSize    = 800,
               pvalueCutoff = 0.05,
               pAdjustMethod = "none",
               keyType       = "kegg")

data<-data.frame(kk2)
write.csv(data ,file="KEGG_results/Male_Jejunum_KEGG_enrichment.csv")
```

```{r}
tiff(file = 'enrichment_plots/Male_Jejunum_enrichment.tiff', units="in", width=10, height=5, res=1000)
dot<-dotplot(kk2, showCategory=5, split=".sign")+
  scale_color_distiller(palette = "RdYlBu", trans= "reverse")+
  facet_grid(.~.sign)+
  ggtitle("Male Jejunum KEGG Enrichment")+                                                           # Change font size
  theme(strip.text.x = element_text(size = 12))+
  theme(axis.text.y = element_text(size = 12))   
dot
dev.off()
dot
```

```{r}
# Produce the native KEGG plot (PNG)
male_liver <- pathview(gene.data=gene_list, pathway.id="00190", species = kegg_organism)

```